openTSNE: a modular \proglang{Python} library for t-SNE dimensionality reduction and embedding

### 2022-10-19 paper/main.tex
minor changes

Easy access to efficient implementations almost universally correlates with the widespread adoption of novel computational techniques. Despite the many theoretical advancements, most popular t-SNE libraries have been slow to incorporate them into their implementations.
In \proglang{Python}, the most commonly used implementation of the t-SNE algorithm comes from \pkg{scikit-learn}~\citep{pedregosa2011scikit}, which suffers from poor scalability. Other implementations from the \proglang{R}~\citep{team2013r,krijthe2015rtsne} and \proglang{Julia}~\citep{bezanson2017julia,julia_tsne} programming languages suffer from similar drawbacks. \pkg{MulticoreTSNE}~\citep{Ulyanov2016} enjoys better scalability but is likewise limited when applied to data sets containing millions of data points. Additionally, none of these implementations include recently proposed extensions that produce more globally consistent embeddings. \pkg{FIt-SNE}~\citep{linderman2019fast} alleviates these problems and provides bindings to both the \proglang{Python} and \proglang{R} programming languages. However, \pkg{FIt-SNE} is difficult to install and distribute and provides an API largely inconsistent with the \proglang{Python} data science ecosystem. None of these existing implementations provide solutions to adding new samples into existing embeddings.
To alleviate these problems, we present \opentsne, our open-source \proglang{Python} implementation of the t-SNE algorithm, complete with its recently proposed extensions. \opentsne is easy to install and includes precompiled binaries available through popular \proglang{Python} package managers. It provides a familiar API, making it suitable as a drop-in replacement for existing t-SNE implementations.

@@

\subsection{Speed}

>>>>

Easy access to efficient implementations almost universally correlates with the widespread adoption of novel computational techniques. Despite the many recent theoretical advancements, popular t-SNE libraries have slowly incorporated them into their implementations.
In \proglang{Python}, the most commonly used implementation of the t-SNE algorithm comes from \pkg{scikit-learn}~\citep{pedregosa2011scikit}, which is nicely integrated with \proglang{Python} data science environment but suffers from poor scalability. Other implementations from the \proglang{R}~\citep{team2013r,krijthe2015rtsne} and \proglang{Julia}~\citep{bezanson2017julia,julia_tsne} programming languages also scale poorly (see Sec.~\ref{s:speed}). \pkg{MulticoreTSNE}~\citep{Ulyanov2016} enjoys better scalability but is likewise limited when applied to larger data sets. Additionally, none of these implementations include recently proposed extensions that can result in more globally consistent embeddings. \pkg{FIt-SNE}~\citep{linderman2019fast} alleviates these problems and provides bindings to both the \proglang{Python} and \proglang{R} programming languages. However, \pkg{FIt-SNE} is difficult to install and distribute and provides an API largely inconsistent with the \proglang{Python} data science ecosystem. None of the existing t-SNE implementations support adding new samples into existing embeddings. To alleviate all these problems, and to support newly proposed additions to t-SNE algorithms, we here present \opentsne, an open-source \proglang{Python} implementation of the t-SNE algorithm. \opentsne is easy to install and includes precompiled binaries available through popular \proglang{Python} package managers. It provides a familiar API, making it suitable as a drop-in replacement for existing t-SNE implementations.

@@

\subsection{Speed\label{s:speed}}
